---
title: "UG-Trust Behavioral Analyses"
output: pdf_document
date: "2026-02-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data <- read.table("sublist_n132.txt", header = FALSE)
colnames(data) <- "sub_id"




# contains fairness sensitivity slopes
fairness_data = read.csv('slope_df.csv')

# contains trial-level information on accept/rejection
behavioral_data = read.csv('behavioral_data.csv')
library(tidyverse)
percent_accept_data <- behavioral_data %>%
  group_by(sub_id) %>%
  summarise(
    sum_accept = sum(accept, na.rm = TRUE),
    total_trials = sum(!is.na(accept)),
    percent_accept = (sum_accept / total_trials) * 100,
    
    # Calculate percent_accept_social only for "social" trials
    sum_accept_social = sum(accept[condition == "social"], na.rm = TRUE),
    total_social = sum(condition == "social" & !is.na(accept)),
    percent_accept_social = (sum_accept_social / total_social) * 100,
    
    # Calculate percent_accept_nonsocial only for "nonsocial" trials
    sum_accept_nonsocial = sum(accept[condition == "nonsocial"], na.rm = TRUE),
    total_nonsocial = sum(condition == "nonsocial" & !is.na(accept)),
    percent_accept_nonsocial = (sum_accept_nonsocial / total_nonsocial) * 100
  )



# continue combining data frames
data <- merge(data, fairness_data,
                     by.x = "sub_id",  # column from 'data'
                     by.y = "sub_id",         # column from 'fairness_data'
                     all.x = TRUE)   
data = merge(data, percent_accept_data,
             by.x = 'sub_id',
             by.y = 'sub_id',
             all.x = TRUE)




demographics <- behavioral_data %>%
  select(sub_id, sub_age, sub_gender) %>%
  distinct() %>%
  mutate(
    male = ifelse(sub_gender == "Male", 1, 0)
  ) %>%
  select(sub_id, sub_age, male)   # drop original sub_gender

data <- merge(data, demographics,
              by = "sub_id",
              all.x = TRUE)
```

```{r}

# read subject list
subs <- scan("sublist_n132.txt", what = character())

# initialize output data frame
out_AIns <- data.frame(
  sub_id = subs,
  trust_AIns = NA_real_,
  ugr_AIns   = NA_real_,
  stringsAsFactors = FALSE
)

# helper function to read the correct file
read_meants <- function(folder, sub) {
  l2_file  <- file.path(folder, paste0("sub-", sub, "_L2_AIns_meants.txt"))
  run1_file <- file.path(folder, paste0("sub-", sub, "_run-1_AIns_meants.txt"))
  
  if (file.exists(l2_file)) {
    as.numeric(scan(l2_file, what = numeric(), quiet = TRUE))
  } else {
    as.numeric(scan(run1_file, what = numeric(), quiet = TRUE))
  }
}

# loop over subjects
for (i in seq_along(subs)) {
  s <- subs[i]
  
  out_AIns$trust_AIns[i] <- read_meants("AIns_trust_meants", s)
  out_AIns$ugr_AIns[i]   <- read_meants("AIns_ugr_meants", s)
}

data <- merge(data, out_AIns,
              by = "sub_id",
              all.x = TRUE)





# initialize output data frame
out_AIns_social <- data.frame(
  sub_id = subs,
  trust_AIns_social = NA_real_,
  ugr_AIns_social   = NA_real_,
  stringsAsFactors = FALSE
)

# helper function to read the correct file
read_meants <- function(folder, sub) {
  l2_file  <- file.path(folder, paste0("sub-", sub, "_L2_AIns_social_meants.txt"))
  run1_file <- file.path(folder, paste0("sub-", sub, "_run-1_AIns_social_meants.txt"))
  
  if (file.exists(l2_file)) {
    as.numeric(scan(l2_file, what = numeric(), quiet = TRUE))
  } else {
    as.numeric(scan(run1_file, what = numeric(), quiet = TRUE))
  }
}

# loop over subjects
for (i in seq_along(subs)) {
  s <- subs[i]
  
  out_AIns_social$trust_AIns_social[i] <- read_meants("AIns_trust_social_meants", s)
  out_AIns_social$ugr_AIns_social[i]   <- read_meants("AIns_ugr_social_meants", s)
}


data <- merge(data, out_AIns_social,
              by = "sub_id",
              all.x = TRUE)









# initialize output data frame
out_dACC <- data.frame(
  sub_id = subs,
  trust_dACC = NA_real_,
  trust_dACC_social = NA_real_,
  ugr_dACC = NA_real_,
  ugr_dACC_social = NA_real_,
  stringsAsFactors = FALSE
)

# helper function to read the correct file
read_dACC_meants <- function(folder, sub) {
  l2_file  <- file.path(folder, paste0("sub-", sub, "_L2_dACC_meants.txt"))
  run1_file <- file.path(folder, paste0("sub-", sub, "_run-1_dACC_meants.txt"))
  
  if (file.exists(l2_file)) {
    as.numeric(scan(l2_file, what = numeric(), quiet = TRUE))
  } else {
    as.numeric(scan(run1_file, what = numeric(), quiet = TRUE))
  }
}

# helper function to read the correct file
read_dACC_social_meants <- function(folder, sub) {
  l2_file  <- file.path(folder, paste0("sub-", sub, "_L2_dACC_social_meants.txt"))
  run1_file <- file.path(folder, paste0("sub-", sub, "_run-1_dACC_social_meants.txt"))
  
  if (file.exists(l2_file)) {
    as.numeric(scan(l2_file, what = numeric(), quiet = TRUE))
  } else {
    as.numeric(scan(run1_file, what = numeric(), quiet = TRUE))
  }
}

# loop over subjects
for (i in seq_along(subs)) {
  s <- subs[i]
  
  out_dACC$trust_dACC[i] <- read_dACC_meants("dACC_trust_meants", s)
  out_dACC$trust_dACC_social[i]   <- read_dACC_social_meants("dACC_trust_social_meants", s)
  out_dACC$ugr_dACC[i] <- read_dACC_meants("dACC_ugr_meants", s)
  out_dACC$ugr_dACC_social[i]   <- read_dACC_social_meants("dACC_ugr_social_meants", s)
}

data <- merge(data, out_dACC,
              by = "sub_id",
              all.x = TRUE)
```
#Get demographics
```{r}
library(dplyr)
library(knitr)
library(kableExtra)
# behavioral_data$sub_id is numeric → convert to character
behavioral_data$sub_id <- as.character(behavioral_data$sub_id)

# Keep only behavioral_data rows where subject is in sub_ids
behavioral_data <- behavioral_data %>%
  filter(sub_id %in% data$sub_id)


# df = your long dataframe
# assumes columns: sub_id, sub_age, sub_gender
demo_tbl <- behavioral_data %>%
  # 1) one row per subject (prevents long-format duplication)
  group_by(sub_id) %>%
  summarise(
    sub_gender = dplyr::first(na.omit(sub_gender)),
    sub_age    = dplyr::first(na.omit(sub_age)),
    .groups = "drop"
  ) %>%
  filter(!is.na(sub_gender), !is.na(sub_age)) %>%
  
  # 2) summarize by gender
  group_by(sub_gender) %>%
  summarise(
    Count    = n(),
    Percent  = 100 * n() / n_distinct(.$sub_id),  # will be overwritten below; see note
    Mean_Age = mean(sub_age),
    SD_Age   = sd(sub_age),
    Min_Age  = min(sub_age),
    Max_Age  = max(sub_age),
    .groups = "drop"
  )

# Fix Percent denominator to be TOTAL N across all subjects (not within-group)
N_total <- behavioral_data %>%
  distinct(sub_id, sub_gender, sub_age) %>%
  filter(!is.na(sub_gender), !is.na(sub_age)) %>%
  nrow()

demo_tbl <- demo_tbl %>%
  mutate(
    Percent  = 100 * Count / N_total,
    Mean_Age = round(Mean_Age, 1),
    SD_Age   = round(SD_Age, 1),
    Min_Age  = round(Min_Age, 2),
    Max_Age  = round(Max_Age, 2),
    Percent  = round(Percent, 1)
  ) %>%
  rename(Gender = sub_gender) %>%
  arrange(match(Gender, c("Female","Male","Non-Binary")))  # optional ordering

# 3) print like your example
kable(
  demo_tbl,
  caption = "Participant Demographics by Gender",
  align = "lrrrrrr"
) %>%
  kable_styling(full_width = FALSE)
```





# Testing H1.1
We hypothesize that elevated dACC responses to unreciprocated trust in the Trust Game will be associated with a greater tendency to reject offers in the UG 

**DV: Sensitivity to fairness sloped; IV: ACC Activation during TG, age, sex**
```{r}
H1.1_slopes = lm(slope_all ~ trust_dACC + sub_age + male, data = data)
summary(H1.1_slopes)
```

**DV: UG Rejection rates; IV: dACC Activation during TG, age, sex**
```{r}
H1.1_rates = lm(percent_accept ~ trust_dACC + sub_age + male, data = data)
summary(H1.1_rates)
```


**DV: Sensitivity to fairness sloped; IV: AIns Activation during TG, age, sex**
```{r}
H1.1_slopes = lm(slope_all ~ trust_AIns + sub_age + male, data = data)
summary(H1.1_slopes)

H1.1_slopes = lm(slope_all ~ ugr_AIns + sub_age + male, data = data)
summary(H1.1_slopes)
```

**DV: UG Rejection rates; IV: AIns Activation during TG, age, sex**
```{r}
H1.1_rates = lm(percent_accept ~ trust_AIns + sub_age + male, data = data)
summary(H1.1_rates)

H1.1_rates = lm(percent_accept ~ ugr_AIns + sub_age + male, data = data)
summary(H1.1_rates)
```



```{r}
H1.1_slopes = lm(slope_all ~ ugr_AIns + sub_age + male, data = data)
summary(H1.1_slopes)
```

```{r}
H1.1_rates = lm(percent_accept ~ ugr_AIns + sub_age + male, data = data)
summary(H1.1_rates)
```

```{r}
H1.2_slopes = lm(slope_social ~ ugr_AIns_social + sub_age + male, data = data)
summary(H1.2_slopes)
```

```{r}
H1.2_rates = lm(percent_accept_social ~ ugr_AIns_social + sub_age + male, data = data)
summary(H1.2_rates)
```



#ANOVA Results
```{r}
library(tidyverse)
library(ez)

### ----------------------------------------------------------
### STEP 1 — CLEAN subject IDs in sub_ids
### ----------------------------------------------------------


# behavioral_data$sub_id is numeric → convert to character
behavioral_data$sub_id <- as.character(behavioral_data$sub_id)

# Keep only behavioral_data rows where subject is in sub_ids
behavioral_data <- behavioral_data %>%
  filter(sub_id %in% data$sub_id)


### ----------------------------------------------------------
### STEP 2 — PREPARE TRIAL-LEVEL ACCEPTANCE VARIABLE
### ----------------------------------------------------------

behavioral_data <- behavioral_data %>%
  mutate(accept_binary = ifelse(accept_reject == "Accept", 1, 0))

### ----------------------------------------------------------
### STEP 3 — CREATE SUBJECT × FAIRNESS × CONDITION ACCEPTANCE RATES
### ----------------------------------------------------------

anova_df <- behavioral_data %>%
  group_by(sub_id, beh_fairness_perc, condition) %>%
  summarise(
    accept_rate = mean(accept_binary, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    # order fairness levels (largest → smallest)
    beh_fairness_perc = factor(
      beh_fairness_perc,
      levels = c("50%", "25%", "10%", "5%")
    ),
    # ensure correct ordering of social condition
    condition = factor(condition, levels = c("social", "nonsocial"))
  )

### ----------------------------------------------------------
### STEP 4 — RUN TWO-WAY REPEATED-MEASURES ANOVA
### ----------------------------------------------------------

anova_results <- ezANOVA(
  data    = anova_df,
  dv      = accept_rate,
  wid     = sub_id,
  within  = .(beh_fairness_perc, condition),
  type    = 3,
  detailed = TRUE
)

anova_results
```





```{r}
library(ggplot2)
library(dplyr)
library(scales)

plot_data <- behavioral_data %>%
  mutate(
    accept_binary = ifelse(accept_reject == "Accept", 1, 0)
  )

plot_df <- plot_data %>%
  group_by(beh_fairness_perc, condition) %>%
  summarise(
    n           = dplyr::n(),
    accept_rate = mean(accept_binary, na.rm = TRUE),
    sd_accept   = sd(accept_binary, na.rm = TRUE),
    se_accept   = sd_accept / sqrt(n),
    .groups     = "drop"
  ) %>%
  mutate(
    condition = factor(condition, levels = c("social", "nonsocial")),
    ymin = pmax(0, accept_rate - se_accept),
    ymax = pmin(1, accept_rate + se_accept),
    beh_fairness_perc = factor(
      beh_fairness_perc,
      levels = c("5%", "10%", "25%", "50%")
    )
  )

pd <- position_dodge(width = 0.75)

p <- ggplot(plot_df, aes(x = beh_fairness_perc, y = accept_rate, fill = condition)) +
  geom_col(
  width = 0.7,
  position = pd,
  linewidth = 0
  ) +
  geom_errorbar(
    aes(ymin = ymin, ymax = ymax, group = condition),
    width = 0.15,
    position = pd,
    linewidth = 0.3
  ) +
  scale_y_continuous(
    name   = "Acceptance Rate",
    limits = c(0, 1),
    breaks = c(0, 0.25, 0.50, 0.75, 1.00),
    labels = label_percent(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_x_discrete(
    name = "Offer Fairness (% of endowment)"
  ) +
  scale_fill_manual(
    values = c("social" = "#E65D2C", "nonsocial" = "#0870B7"),
    breaks = c("social", "nonsocial"),
    labels = c("Social", "Nonsocial")
  ) +
  theme_classic(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    axis.ticks = element_line(linewidth = 0.4),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 12),
    legend.title = element_blank()
  )

p

ggsave(
  "acceptance_rate_plot.tiff",
  plot = p,
  width = 6,
  height = 5,
  dpi = 600
)

ggsave("acceptance_rate_plot_svg.svg", plot = p, width = 6, height = 5)
```


# Testing H1.2
We hypothesize that elevated dACC responses to unreciprocated trust in the Trust Game will be associated with a greater tendency to reject offers in the UG *FOR SOCIAL PARTNER*

**DV: Sensitivity to fairness slope during social UG trials; IV: dACC Activation during TG with social partner, age, sex**
```{r}
H1.2_slopes = lm(slope_social ~ trust_dACC_social + sub_age + male, data = data)
summary(H1.2_slopes)
```

**DV: UG Rejection rates during social trials; IV: dACC Activation during TG with social partner, age, sex**
```{r}
H1.2_rates = lm(percent_accept_social ~ trust_dACC_social + sub_age + male, data = data)
summary(H1.2_rates)
```


**DV: Sensitivity to fairness slope during social UG trials; IV: AIns Activation during TG with social partner, age, sex**
```{r}
H1.2_slopes = lm(slope_social ~ trust_AIns_social + sub_age + male, data = data)
summary(H1.2_slopes)

H1.2_slopes = lm(slope_social ~ ugr_AIns_social + sub_age + male, data = data)
summary(H1.2_slopes)
```

**DV: UG Rejection rates during social trials; IV: AIns Activation during TG with social partner, age, sex**
```{r}
H1.2_rates = lm(percent_accept_social ~ trust_AIns_social + sub_age + male, data = data)
summary(H1.2_rates)

H1.2_rates = lm(percent_accept_social ~ ugr_AIns_social + sub_age + male, data = data)
summary(H1.2_rates)
```






# Testing H2.1 and H2.2
```{r}
H2.1 = lm(ugr_dACC ~ trust_dACC + sub_age + male, data = data)
summary(H2.1)

H2.2 = lm(ugr_dACC_social ~ trust_dACC_social + sub_age + male, data = data)
summary(H2.2)

H2.1 = lm(ugr_AIns ~ trust_AIns, data = data)
summary(H2.1)

H2.2 = lm(ugr_AIns_social ~ trust_AIns_social, data = data)
summary(H2.2)
```

```{r}
library(tidyverse)

plot_data <- data %>%
  drop_na(ugr_AIns, trust_AIns, sub_age, male) %>%
  mutate(
    ugr_AIns_adj = resid(lm(ugr_AIns ~ sub_age + male, data = .)) + 
      mean(ugr_AIns, na.rm = TRUE),
    
    trust_AIns_adj = resid(lm(trust_AIns ~ sub_age + male, data = .)) + 
      mean(trust_AIns, na.rm = TRUE)
  )

p = ggplot(plot_data, aes(x = trust_AIns_adj, y = ugr_AIns_adj)) +
  geom_point(size = 2.5, alpha = 0.75) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    linewidth = 1,
    color = "#F5E911"
  ) +
  labs(
    x = "TG aIns Response to Defected Trust (z)",
    y = "UG aIns Response to Unfairness (z)",
    title = "Association Between TG and UG aIns Response to Norm Violations"
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 18)
  )

ggsave("crosstask_aIns_plot.svg", plot = p, width = 6, height = 5)
```

